---
title: Survey Report of the Liberal Party Supporters in the University-Rosedale Riding
  District with Data Simulation and Analysis
output:
  html_document:
    df_print: paged
---
Summary
====================================   
This project aims at surveying the background information of Liberal Party supporters in the University-Rosedale riding district. A number of 2000 randomly selected participants in the region will receive an online questionnaire through email or browser implantation. The sampling process uses a technique called Simple Random Sampling without replacement to make sure the participants are chosen without bias and are good representatives for the population in the region. Assistant methods of locating respondents include but not is limited to postal code tracking. This step is crucial as it will make sure the participant is our desired target. The respondents will be reached online with full anonymity through email and web-links. All answers will be kept confidential and collected through @Survey_Monkey. A simulation via @R_studio will perform to predict the distribution of the answers to the questionnaires. Data distribution from previous elections and population census are referenced to a better approach to reality. The simulated result is thoroughly discussed as well in contemplation to predict the outcome of the survey. Regard to the fact that only the participants who support the Liberal are interested in the project, some samples will be discarded to avoid non-responder or biased answers even though this will cause excess expense. Although this will cause excess expense, it is an essential step to avoid non-response or biased answers. In the simulation, the middle class participants form the majority of Liberal supporters, whereas the people who makes more than 150 thousand CAD per year are rare in the group. From the perspective of educational background, a predominant amount(about 70%) of respondents who support liberal party are well-educated(received at least some post-secondary education). These findings are fair estimates for predicting the background of Liberal supporters in the University-Rosedale riding district, which help the Liberal Party staffs better plan for their future policy and campaign. As discussed, this project mainly focuses on Liberal supporters in a certain riding district. Therefore, the simulation that referenced the population census data of Ontario might be somewhat biased. After-all, the excess survey data can be utilized in future researches such as analyzing the population structure of supporters of the ‘enemy parties’. Web-link to the survey and screenshots of the questionnaire are attached in the appendix.
&nbsp;    
&nbsp;    
&nbsp;    
                       
Introduction
====================================   
Each seat in the House of Common represents the voice of one specific riding district in the broad land of Canada. Understanding the people in each riding district is absolutely critical for winning more votes. In this survey project, the primary target is to collect and analyze the background information of the voters who support the Liberal Party in the University-Rosedale riding district. Respondents will be randomly sampled and reached via online surveys with techniques that ensure their unbiasedness and representativeness. Online questionnaires will be designed and delivered online via @Survey_Monkey(survey link and screenshots are attached in the appendix). Data in various aspects will be collected and the primary focus would be participants’ educational background and annual income range. Detailed survey methodologies will be thoroughly discussed in the coming section, followed by cost estimates. Common concerns such as non-responses and participant privacy issues are discussed and well managed respectively. A data simulation process is carried out in order to estimate the outcome of the survey via @R_studio. Graphs of simulated responses to each survey question are constructed and discussed in the result session.

&nbsp;    
&nbsp;    
&nbsp;    
              
Sampling Strategy and Survey Methodology   
====================================   
The population of this survey is all the voters who live in the University-Rosedale riding district. The Data frame will be constructed using internet strategies to target email users who have the postal codes that are in the desired region. A number of 2000 samples will be drawn randomly to complete this survey using Simple Random Sampling(SRS) without replacement to make the sample as unbiased and representative as possible. The University-Rosedale riding district has a population of 104,000, which makes it somewhat hard to obtain the entire sample frame. However,using advanced online tracking technology, it is feasible for us to construct a relatively complete sampling frame.  
   
Note that, since we are only interested in people who support the Liberal Party in this particular research, we will screen the sample and focus on the people who support the Liberal Party. The rest of the surveying results can be utilized in other related researches. This is costly but necessary step as regards minimizing non-response bias.   
   
To reach the desired respondents, the major surveying technique employed is an online questionnaire containing 7 questions via @Survey_Monkey (link to the survey and screenshots are included in the Appendix). The surveys will be delivered online, tracking internet users who have a postal code in the University-Rosedale riding district. Internet IP addresses tracking can be supplementary. For the sake of ‘none replacement’, one specific device under the same IP address can only fill out the survey once. Questionnaires will be delivered to the respondent through email, internet browser implantation, etc.. Data will be collected directly from the @Survey_Monkey website for analysis.   
&nbsp;    
&nbsp;    

Estimated Cost
====================================   
The estimated cost that survey monkey shows is 12800 in CAD. However, the actual estimated cost would be a bit higher since more precise delivery for survey in a specific distraction is needed. We chose to deliver all those 2,000 surveys online in order to reduce the cost as much as possible. Although, we only used 512 people’s data at the end, all the other data can be applied on other areas. In other words, no single collected data will be wasted in that case.   
&nbsp;    
&nbsp;    
&nbsp;         
       
                          
Plans for Non-Responders and Non-response Bias  
====================================   
To minimize the effect of non-responses, efforts must be put into survey designing. The questionnaire is concise and straight forward. In the meantime, it is considerable to all groups of people so that none would feel offended or excluded throughout the survey (eg. the LGTBQ group). From the political aspect, although the targeted group is Liberal supporters, all respondents’ data will be collected regardless of their political standings. These steps will be costly but the excess data will be valuable in other researches. The non-respond estimate given by @Survey_Monkey is 20% for the survey, which is within an acceptable range for our analysis.
&nbsp;    
&nbsp;    


                                    
Respondent Privacy Concerns
====================================   
All the survey answers collected will be kept confidential and only used for this analysis. During the survey, respondents will keep full anonymity. That means, no questions related to the respondent's name, address, or email address will be asked. In addition to that, personal information such as age and income will be answered in a range instead of a specific number to minimize privacy concerns. 
&nbsp;    
&nbsp;    
     
        
               
Potential Weaknesses and Efforts for Amendment     
====================================   
There are a few inevitable biases present in the survey methodology and data simulation. In terms of sampling bias, the major challenge of SRS is getting the complete frame of the population. People who don’t have stable access to the internet or those who have issues using electronic devices might be excluded from the sampling process.    
    
Given the fact that all surveys are done online, limited methods of distribution can lead to biased data. Besides, since all participants do the survey anonymously, it is hard to make sure whether the person filling out the survey is our target. To minimize the impact of these issues, Postal code is an efficient discriminator chosen for screening the respondent’s voting district. The questionnaire is designed as simple as possible to avoid non-responses.     
&nbsp;   
&nbsp;        
            
                   
Data Simulation Results and Discussion
====================================   
The result of data simulation for the survey is presented and discussed in this section. The distribution of answers for each question is simulated referring to real-world data. The distribution ratio for each answer is referenced from past years’ election and population census data from @election.ca, @ontario.ca, and other statistical websites. Out of the 2000 questionnaires distributed, an expectation of 85% will be filled out and returned. Meanwhile, the question that asks about the participants’ political party standing will be the screening question to filter out the non-supporters(note that, ONLY Liberal party supporters are targeted in the research). The filtered samples will each have a unique answer to each question in the survey, following the distribution probability as mention above. Using data from previous years as a template can lead to certain errors due to certain population structure differentiation from the population census in 2016. However, few adjustments are made to better approach real-world situations. For example, in the gender question, the choice of “other gender” will appear slightly more frequent since the Liberal Party’s ideology is more tolerant of the LGTBQ group.
&nbsp;    
         
            
All data available at : https://github.com/Bowen-Han/The-Background-of-Liberal-Voters
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(4736)
#Total 2000 surveys
total<-2000
#Expect for 80% completion.
expect<-total*0.8
#step1. Set all survey choices.
#The proportion of voting each Parties.
liberal<-expect*0.32
conservative <-expect*0.34
bloc <- expect*0.09
NDP <- expect*0.07
Green <-expect*0.008
others <- expect-liberal-conservative-bloc-NDP-Green
party <- c("Liberal", "Conservative",  "Bloc Québécois","New Democratic Party","Green Party of Canada","Others")
gender <- c("Male", "Female", "Other", "Prefer not to say")
age_group<-c("18-24", "25-34", "35-44", "45-54","55-64","65-74","75+")
annual_income <-c("A.Under 5k","B.5k-9.9k","C.10k-14.9k","D.15k-19.9k","E.20k-24.9k","F.25k-34.9k","G.35k-49.9k","H.50k-74.9k","I.75k-99.9k","J.100k-149.9k","K.150k-199.9k","L.200k and more")
edu_background <-c("1.Before High school","2.High school","3.Postsecondary certificate school","4.Bachelor's degree","5.Above Bachelor's degree","Others")
Immigrantion_status <-c("Yes","No")
primary_language <- c("English","French","Others")

#step2.Set the probability of each choise.
#The proportion of each gender vote for Libearls.
gender.liberal<-sample(gender,liberal,replace = TRUE,prob = c(0.48,0.51,0.008,0.002))
#The proportion of each age group vote for Libearls.
age.liberal<-sample(age_group,liberal,replace=TRUE,prob=c(0.0933,0.1448,0.1562,0.1639,0.1815,0.2092,0.0512))
#The proportion of each income group vote for Libearls.
income.liberal <- sample(annual_income,liberal,replace = TRUE,prob = c(0.061,0.053,0.075,0.079,0.086,0.125,0.163,0.172,0.09,0.063,0.016,0.017))
#The proportion of people with different educational level vote for Libearls.
edu.liberal <- sample(edu_background,liberal,replace = TRUE,prob = c(0.1090,0.1993,0.3555,0.2192,0.1169,0.0001))
#The proportion of locals and immigrants vote for Libearls.
immigrantion.liberal <- sample(Immigrantion_status,liberal,replace = TRUE,prob = c(0.22,0.78))
#The proportion of each language group vote for Libearls.
language.liberal <- sample(primary_language,liberal,replace = TRUE,prob = c(0.682,0.04,0.279))
#Then generate dataframe from all six vectors.
riding_data<-data.frame(Gender=gender.liberal,
                        Party = "Liberal",
                        Annual_income = income.liberal,
                        Education_Background = edu.liberal,
                        Immigration = immigrantion.liberal,
                        Mother_language = language.liberal,
                        Age=age.liberal)
```
     
     
```{r,echo=FALSE, message=FALSE}
#Re-group the dataset by Gender.
sex <- riding_data %>%
  group_by(`Gender`) %>%
  summarise(counts = n())
#Creat bar chart about gender.
sex1 <- ggplot(sex,aes(x=Gender,y=counts)) +
  geom_bar(fill=("light blue"),stat = 'identity') +
  geom_text(aes(label=counts)) +
  theme(axis.text.x = element_text(angle=0),
        plot.title=element_text(colour='Dark blue', hjust=0.5,face='bold'))+
  labs(title = "The Gender Distribution of Liberal Supporters-1.1")
sex1

```
Out of a total of 1,600 surveys that were collected, 512 of them supported the Liberal Party. According to figure 1.1- ‘The Gender Distribution of Liberal Supporters', 262 are females and 242 are males. There is no significant difference between sexes, meaning that the Liberal Party does not have any special policies or properties that attract one specific gender. Moreover, 6 participants who responded as "other gender" which possibly suggests that the Liberals' ideological policies benefit the minority groups.
```{r,echo=FALSE, message=FALSE}
#calculate the proportion of each age group.
age <- riding_data %>%
  group_by(`Age`) %>%
  summarise(counts = n())
Prop_age <- age %>%
  mutate(prop = round(counts*100/sum(counts), 1))
#Set the position of label on chart.
cumulative = cumsum(Prop_age$counts)
midpoint = cumulative - Prop_age$counts / 2
#Drawing pie chart according age proportion.
ggplot(Prop_age, aes(x = "", y = prop, fill = `Age`)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
   labs(title = "Distribution of Age Group of Liberal Supporters-2.1")+
  theme_void()+
  geom_text(stat="identity",aes(label = scales::percent(`prop`/100)), size=3, position=position_stack(vjust = 0.5))
#Drawing the distribution of age group by bar chart.
age1 <- ggplot(age,aes(x=Age,y=counts)) +
  geom_bar(fill=("light yellow"),stat = 'identity') +
  geom_text(aes(label=counts)) +
  theme(axis.text.x = element_text(angle=0),
        plot.title=element_text(colour='Dark green', hjust=0.5,face='bold'))+
  labs(title = "Distribution of Age Group of Liberal Supporters-2.2")
age1
```
Figure 2.1 and 2.2 - ‘Distribution of Age Group of the Liberal Supporters’ demonstrate the distribution of the age groups that Supporters the Liberal Party. A majority of 120 supporters aged between 65 and 74 years old voted for the Liberals, which occupied for 23.4% of the total respondents. Coming next are respondents in age groups “55 to 64” and “45 to 54” accounting for 17.6% and 16.8% respectively. The data in this section is simulated with flaws since the Liberal Party is probably not the most popular political party among elders. 
```{r,echo=FALSE, message=FALSE}
#Set new dataset for annual income.
income <- riding_data %>%
  group_by(`Annual_income`) %>%
  summarise(counts = n())
#Creat the new bar chart to show the distribution of income.
income1 <- ggplot(income,aes(x=Annual_income,y=counts)) +
  geom_bar(fill=("light yellow"),stat = 'identity') +
  geom_text(aes(label=counts)) +
  theme(axis.text.x = element_text(angle=90),
        plot.title=element_text(colour='Dark green', hjust=0.5,face='bold'))+
  labs(title = "Distribution of Annual Income in CAD of Liberal Supporters-3.1")
income1
#Calculate the proportion of each income group.
Prop_income <- income %>%
  mutate(prop = round(counts*100/sum(counts), 1))
#Set the position of label on pie chart.
cumulative = cumsum(Prop_income$counts)
midpoint = cumulative - Prop_income$counts / 2
#Darwing the radar chart for income groups.
ggplot(data = riding_data) +

  geom_bar(

    mapping = aes(x = Annual_income, fill = Annual_income),

    show.legend = FALSE,

    width = 1

  ) +

  theme(aspect.ratio = 1) +
  theme(legend.title = element_blank(), legend.position = "top") +
  labs(title = "Distribution of Annual Income in CAD of Liberal Supporters-3.2")+
  labs(x = NULL, y = NULL) +
  coord_polar()
#Show the proportion that calculated by pie chart.
ggplot(Prop_income, aes(x = "", y = prop, fill = `Annual_income`)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
   labs(title = "Distribution of Annual Income in CAD of Liberal Supporters-3.3")+
  theme_void()+
  geom_text(stat="identity",aes(label = scales::percent(`prop`/100)), size=3, position=position_stack(vjust = 0.5))
```
Figure 3.1-3.3 ‘Distribution of Income in CAD of Liberal Supporters’ illustrates the detailed distribution of Liberal supporters’ income ranges. Their annual income is mainly in $35,000 to $49,999 and $50,000 to $74,999, each contain 75 and 82 people respectively. The income groups $35,000 to $49,999 and $50,000 to $74,999 in Canada are representatives of the middle class and the summation of these two gropes’ proportion in total Liberals votes is 30.6%. This result states that Liberals rely mainly on the middle class’s support since they attribute to almost 1/3 of the total. On the other hand, participants who earn more than $150,000 per year occupies only 3.2% of total Liberals votes. (Although those two groups would make much more money compared to all the others, the rare quantity makes them indecisive for party election.) Therefore, if the Liberal Party wants to draw more attention in the next election, they might need to focus more on the middle class.
```{r,echo=FALSE, message=FALSE}
#Creat a new dataset according to the Education_Background.
edu <- riding_data %>%
  group_by(`Education_Background`) %>%
  summarise(counts = n())
#Drawing bar chart for the new dataset.
edu1 <- ggplot(edu,aes(x=Education_Background,y=counts)) +
  geom_bar(fill=("Dark Green"),stat = 'identity') +
  geom_text(aes(label=counts)) +
  theme(axis.text.x = element_text(angle=90),
        plot.title=element_text(colour='light green', hjust=0.5,face='bold'))+
  labs(title = "Distribution of Educational Level of Liberal Supporters-4.1")
edu1

#Calculate the proportion of each level of education.
Prop_edu <- edu %>%
  mutate(prop = round(counts*100/sum(counts), 1))
cumulative = cumsum(Prop_edu$counts)
midpoint = cumulative - Prop_edu$counts / 2
#Drawing the pie chart for each educational level.
ggplot(Prop_edu, aes(x = "", y = prop, fill = `Education_Background`)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
   labs(title = "Distribution of Educational Level of Liberal Supporters-4.2")+
  theme_void()+
  geom_text(stat="identity",aes(label = scales::percent(`prop`/100)), size=3, position=position_stack(vjust = 0.5))
```
Figure 4.1-4.2 "Distribution of Educational Level of Liberal Supporters" indicates the educational backgrounds of the participants. 195 people graduated from certificated postsecondary schools, which occupies 38.1% of the total Liberals votes, accounting for the predominant proportion. Notably, according to the overall educational background of Ontario, people with a Bachelor’s degree (or higher) are only a minority compared to people who got less education. Therefore, 21.5 and 11.5 percent from the group “Bachelor’s degree” and group “Master degree or higher” respectively is considered high, attributing to a remarkable proportion of Liberals votes. This shows that the Liberal Party is more popular among the well-educated population. The reason might be Liberals policy is easier for those groups of people to understand and this feature needs to be kept by Liberals for taking advantage.  
```{r,echo=FALSE, message=FALSE}
#Creat a new dataset according to the immigration status.
immigration <- riding_data %>%
  group_by(`Immigration`) %>%
  summarise(counts = n())
#Calculate the proportion of immigrants and local citizens.
Prop_immigration <- immigration %>%
  mutate(prop = round(counts*100/sum(counts), 1))
#Set the postion of label on pie chart.
cumulative = cumsum(Prop_immigration$counts)
midpoint = cumulative - Prop_immigration$counts / 2
#Creat the pie chart about immigration status.
ggplot(Prop_immigration, aes(x = "", y = prop, fill = `Immigration`)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
   labs(title = "Distribution of Immigration Status of Liberal Supporters-5.1")+
  theme_void()+
  geom_text(stat="identity",aes(label = scales::percent(`prop`/100)), size=3, position=position_stack(vjust = 0.5))
```
Figure 5.1 shows the percentage of immigrant population among Liberal supporters. The number of immigrants in Canada is 313.6 thousand in 2019, which is a small portion compared to the population in Canada. However, about one-fifth of Liberal supporters are immigrants according to the survey. Such a high proportion indicates how the Liberal's ideology favors immigrants.
```{r,echo=FALSE, message=FALSE}
#Creat a new dataset according to the primary language.
language <- riding_data %>%
  group_by(`Mother_language`) %>%
  summarise(counts = n())
#Drawing bar chart to show the distribution of using language.
language1 <- ggplot(language,aes(x=Mother_language,y=counts)) +
  geom_bar(fill=("Orange"),stat = 'identity') +
  geom_text(aes(label=counts)) +
  theme(axis.text.x = element_text(angle=0),
        plot.title=element_text(colour=' red', hjust=0.5,face='bold'))+
  labs(title = "Distribution of Primary Language of Liberal Supporters-6.1")
language1
```
Figure 6.1 reflects that the anglophones are the absolute majority who support the Liberal Party. Contrarily, francophones count only 20 out of 512 valid votes. A considerable amount (145) of people who have other languages as primary languages favor the Liberal Party as well.
&nbsp;    
&nbsp;    
&nbsp;     

Reference
----------------------

https://www.elections.ca/res/rep/off/ovr2019app/home.html#17     
https://www150.statcan.gc.ca/n1/pub/89-503-x/2010001/article/11475-eng.htm     
https://www.elections.ca/content.aspx?section=res&dir=rec/part/estim/42ge&document=p1&lang=e#a    
https://www.statista.com/statistics/444858/canada-resident-population-by-gender-and-age-group/     
https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/dv-vd/inc-rev/index-eng.cfm    
https://www.statista.com/statistics/484838/income-distribution-in-canada-by-income-level/     
https://www.statista.com/statistics/606823/population-of-ontario-by-educational-attainment/     
https://www12.statcan.gc.ca/census-recensement/2016/as-sa/fogs-spg/Facts-can-eng.cfm?Lang=Eng&GK=CAN&GC=01&TOPIC=7     
https://www.fin.gov.on.ca/en/economy/demographics/census/cenhi16-6.html     

    
